On centrality of even orthogonal $\Kt$

### 2016-04-27 paper.tex
Fixed statement of Theorem 3

\title{On centrality of orthogonal $\Kt$}
\keywords {Steinberg groups, $K_2$-functor. {\em Mathematical Subject Classification (2010):} 19C09}

@@

\section{Yet another presentation for $\St(4, R)$.} \label{sec:yap}
The key ingredient in Tulenbaev's proof of the local-global principle is the construction of the homomorphism $T$ which fits into the following commutative diagram.
$$\xymatrix{\St(n,\,R\ltimes XR_a[X],\,XR_a[X]) \ar@{^{(}->}[r] \ar[d] & \St(n, R \ltimes XR_a[X]) \ar[d]_-{\lambda_a^*} \\
             \St(n,\,R_a[X],\,XR_a[X]) \ar@{-->}[ru]_-{T} \ar@{^{(}->}[r] & \St(n, R_a[X])}$$
Here $n\geq5$, $a$ is arbitrary nonnilpotent element of $R$ and $\lambda_a^*$ is the morphism induced by the morphism $\lambda_a\colon R \to R_a$ of principal localisation at $a$.
Notice that $XR_a[X]$ is a splitting ideal for both $R_a[X]$ and $R \ltimes XR_a[X]$.
It is clear that the image of $T$ is contained in the image of the top arrow, therefore this implies that the relative Steinberg groups in the left-hand side of the diagram are isomoprhic.
%This $T$ is an inverse to a map induced by localisation, so that constructing $\mathrm T$ may be treated
%as a proof of excision property for Steinberg group in this very particular situation.

@@

In order to construct the arrow $T$ Tulenbaev uses the following presentation for the relative Steinberg group, see~\cite[Proposition~1.6]{Tul}. 
%Assume that $I\trianglelefteq R$ such that a short exact sequence
%\renewcommand{\theequation}{$\dagger$}
\begin{prop}\label{prop:TulPres}
For a splitting ideal $I$ and $n\geq 4$ the group $\St(n,\,R,\,I)$ is isomorphic to the group defined by generators
$$\{X(u,\,v)\mid u\in\E(n,\,R)e_1,\ v\in I^n,\ u^tv=0\}$$ and the following set of relations
\setcounter{equation}{0}
\renewcommand{\theequation}{T\arabic{equation}}
\begin{align}
&X(u,\,v)X(u,\,w)=X(u,\,v+w), \label{add2}\\
&X(u,\,v)X(u',\,v')X(u,\,v)\inv=X(t(u,\,v)u',\,t(v,\,u)\inv v'), \label{conj2}  \\
&X(ur+w,\,v)=X(u,\,vr)X(w,\,v)\,\text{ for }r\in R,\ (u,\,w)\in\Um_{n\times2}(R) \label{add3}.
\end{align}
Moreover, the group $\St(n, R)$ decomposes into the semidirect product $\St(n,\,R, I)\rtimes \St(n,\,R/I)$.
\end{prop}
Here $\Um_{n\times2}(R)$ denotes the set of $n\times2$ \emph{unimodular matrices}, i.\,e. matrices $M$ such that there exists a $2\times n$ matrix $N$ satisfying $NM=\begin{pmatrix}1&0\\0&1\end{pmatrix}$.
In the definition of the semidirect product the group $\St(n, R/I)$ acts on $\St(n, R, I)$ by conjugation via the splitting map, i.\,e. $g \cdot X(u,\,v) = X(\phi(s(g))u, \phi(s(g))^*v)$.

@@

Define the map $\psi$ by $\psi(x_{ij}(\xi)) = (X(e_i, (\xi - \pi(\xi)e_j),\ x_{ij}(\pi(\xi)))$. 
We have to check that $\psi$ preserves the relations \eqref{add0}--\eqref{ccf2}.
For example, let us check \eqref{ccf1}. 
First of all, recall that in any semidirect product of groups one can compute the commutator of two elements as follows:
$$[(a, b), (c, d)] = (a \cdot {}^bc \cdot {}^{bdb^{-1}}a^{-1} \cdot {}^{[b, d]}c^{-1} , [b, d]).$$
For $\xi\in R$ set $\xi' = \xi - \pi(\xi)$.
Specializing $a = X(e_i, \xi' e_j)$, $b = x_{ij}(\pi(\xi))$, $c = X(e_j, \eta'e_k)$, $d = x_{jk}(\pi(\eta))$ and using the definition of the conjugation action 
one computes the commutator $[\psi(x_{ij}(\xi), \psi(x_{jk}(\eta)]$ as follows:
$$\left(X(e_i, \xi' e_j) \cdot X(e_j + \pi(\xi)e_i, \eta'e_k) \cdot X(e_i, \pi(\eta)\xi' e_k - \xi' e_j) \cdot X(e_j, -\eta'e_k),\ x_{ik}(\pi(\xi\eta))\right).$$
After applying additivity relations \eqref{add2}, \eqref{add3} and moving the factor $X(e_i, -\xi' e_j)$ to the left hand side of the formula by means of relation \eqref{conj2}
one gets the following expression
$$(X(e_j, \eta'e_k) \cdot X(e_i, (\xi' \eta' + \pi(\xi)\eta' + \pi(\eta)\xi')e_k \cdot X(e_j, -\eta'e_k),\ x_{ik}(\pi(\xi\eta)).$$
Since $\xi'\eta' + \pi(\xi)\eta' + \pi(\eta)\xi' = \xi\eta - \pi(\xi\eta)$ the above expression clearly simplifies to $\psi(x_{ik}(\xi\eta))$, as claimed.
\end{proof}

@@

In order to construct the arrow $\kappa$ we use the presentation from \cref{prop:TulPres}.
We set $$\kappa(X(u,v)) = F(u,v),\ u\in \E(n, R)e_1,\ v\in I^n.$$ 
In view of \cref{rk:T123'} it suffices to check that $F(u,v)$ satisfy the relations \eqref{add2}, \eqref{conj2}, \eqref{add3'}.
The validity relations \eqref{add2}, \eqref{conj2} is obvious while \eqref{add3'} follows from the following computation:
$$ F(e_i, rae_k) F(e_j, ae_k) = S(rae_i, e_k) S(ae_j, e_k) = S(a(re_i+e_j), e_k) = F(e_ir+e_j, ae_k). \qedhere $$

@@

The following result is a consequence of the above argument.
\begin{lm} \label{lm:map-iota}
 For any $I \trianglelefteq R$ there exists a well-defined homomorphism $$\iota\colon\St^*(n,\,R,\,I)\rightarrow\St(n,\,R, I) \subseteq \St(n, R)$$ 
 satisfying the equalities $\iota (F(u,\,v)) = X(u,\,v)$ and $\iota (S(u,\,v)) = Y(u,\,v)$. \end{lm}
\begin{proof}
We already know that $\iota$ is well-defined. It suffices to check that the image of $\iota$ is contained in $\St(n,\,R,I)$.
Denote the canonical projection $R\epi R/I$ by $\rho$.
Clearly, the induced group morphism $\rho^*\colon\St(n,\,R)\rightarrow\St(n,\,R/I)$ sends $X(u,\,v)$ to $X(\rho(u),\,\rho(v))$ and $Y(u,\,v)$ to $Y(\rho(u),\,\rho(v))$. 
Observe that $\E(n,\,R)\epi\E(n,\,R/I)$ is surjective, therefore for $x\in\E(n,\,R)e_1$ one has $\rho(x)\in\E(n,\,R/I)e_1$. 
From the additivity relations we get that $X(u,\,0)=1=Y(0,\,v)$, hence $\mathrm{Im}(\iota)\subseteq\Ker(\rho^*)=\St(n,\,R,\,I)$, as required. \end{proof}

@@

At this point we have demonstrated for $n\geq4$ and a splitting ideal $I$ that $\St^*(n,\,R,\,I)$ and $\St(n,\,R,\,I)$ are isomorphic.
Now we turn to the main result of this section, namely, we construct a map $$T\colon\St(n,\,R_a[X],\,XR_a[X])\rightarrow\St(n,\,R\ltimes XR_a[X],\,XR_a[X]),$$ for $n\geq4$. 
A local-global principle for Steinberg group and centrality of $\mathrm K_2$ formally follow from existance of this map~\cite[Lem.~15, Lem.~16, proof of Th.~2]{SCh}.

We work in a more general situation.

\begin{tm}
\label{a3map}
Let $B$ be a ring, $I\trianglelefteq B$, $a\in B$ such that $\forall\,x\in I$ there exists a unique $y\in I$ such that $ya=x$ {\rm(}we denote this $y$ by $\frac xa$, elements $\frac x{a^m}$ are also well defined{\rm)}. This requirement is equivalent to say that the principle localisation $\lambda_a\colon I\rightarrow I_a$ of ideal $I$ is an isomorphism. Then there exists a map
$$
\mathrm T\colon\St(n,\,B_a,\,I)\rightarrow\St(n,\,B)
$$
making the diagram
$$
\xymatrix{
\St^*(2n,\,B,\,I)\ar@<-0.0ex>[rr]^{\iota}\ar@<-0.0ex>[d]_{\lambda_{a}^*}&&\St(2n,\,B)\ar@<-0.0ex>[d]^{\lambda_{a}^*}\\
\St^*(2n,\,B_a,\,I)\ar@<-0.0ex>[rr]_{\iota}\ar@{-->}[rru]_{\mathrm T}&&\St(2n,\,B_a)
}
$$
commutative.
\end{tm}

>>>>

\title{On centrality of even orthogonal $\Kt$}
\keywords {Steinberg groups, $K_2$-functor. {\em Mathematical Subject Classification (2010):} 19C09}

@@

\section{Yet another presentation for the relative Steinberg group.} \label{sec:yap}
The key ingredient in Tulenbaev's proof of the local-global principle is the construction of the homomorphism $T$ which fits into the following commutative diagram.
\numberwithin{equation}{section}
\begin{equation} \xymatrix{\St(n,\,R\ltimes XR_a[X],\,XR_a[X]) \ar@{^{(}->}[r] \ar[d] & \St(n, R \ltimes XR_a[X]) \ar[d]_-{\lambda_a^*} \\
             \St(n,\,R_a[X],\,XR_a[X]) \ar@{-->}[ru]_-{T} \ar@{^{(}->}[r] & \St(n, R_a[X])} \label{eq:diagram} \end{equation}
Here $n\geq5$, $a$ is arbitrary nonnilpotent element of $R$ and $\lambda_a^*$ is the morphism induced by the morphism $\lambda_a\colon R \to R_a$ of principal localisation at $a$.
Notice that $XR_a[X]$ is a splitting ideal for both $R_a[X]$ and $R \ltimes XR_a[X]$.
It is clear that the image of $T$ is contained in the image of the top arrow, therefore this implies that the relative Steinberg groups in the left-hand side of the diagram are isomoprhic.
%This $T$ is an inverse to a map induced by localisation, so that constructing $\mathrm T$ may be treated
%as a proof of excision property for Steinberg group in this very particular situation.

@@

In order to construct the arrow $T$ Tulenbaev uses the following presentation for the relative Steinberg group, see~\cite[Proposition~1.6]{Tul}. 
%Assume that $I\trianglelefteq R$ such that a short exact sequence
%\renewcommand{\theequation}{$\dagger$}
\begin{prop}\label{prop:TulPres}
For a splitting ideal $I$ and $n\geq 4$ the group $\St(n,\,R,\,I)$ is isomorphic to the group defined by generators
$$\{X(u,\,v)\mid u\in\E(n,\,R)e_1,\ v\in I^n,\ u^tv=0\}$$ and the following set of relations
\setcounter{equation}{0}
\renewcommand{\theequation}{T\arabic{equation}}
\begin{align}
&X(u,\,v)X(u,\,w)=X(u,\,v+w), \label{add2}\\
&X(u,\,v)X(u',\,v')X(u,\,v)\inv=X(t(u,\,v)u',\,t(v,\,u)\inv v'), \label{conj2}  \\
&X(ur+w,\,v)=X(u,\,vr)X(w,\,v)\,\text{ for }r\in R,\ (u,\,w)\in\Um_{n\times2}(R) \label{add3}.
\end{align}
Moreover, the group $\St(n, R)$ decomposes into the semidirect product $\St(n,\,R, I)\rtimes \St(n,\,R/I)$.
\end{prop}
Here $\Um_{n\times2}(R)$ denotes the set of $n\times2$ \emph{unimodular matrices}, i.\,e. matrices $M$ such that there exists a $2\times n$ matrix $N$ satisfying $NM=\begin{pmatrix}1&0\\0&1\end{pmatrix}$.
In the definition of the semidirect product the group $\St(n, R/I)$ acts on $\St(n, R, I)$ by conjugation via the splitting map, i.\,e. ${}^g X(u,\,v) = X(\phi(s(g))u, \phi(s(g))^*v)$ for $g \in \St(n, R/I)$.

@@

Define the map $\psi$ by $\psi(x_{ij}(\xi)) = (X(e_i, (\xi - \pi(\xi)e_j),\ x_{ij}(\pi(\xi)))$. 
We have to check that $\psi$ preserves the relations \eqref{add0}--\eqref{ccf2}.
For example, let us check \eqref{ccf1}. 
First of all, recall that in any semidirect product of groups one can compute the commutator of two elements as follows:
$$[(a, b), (c, d)] = (a \cdot {}^bc \cdot {}^{bdb^{-1}}a^{-1} \cdot {}^{[b, d]}c^{-1} , [b, d]).$$
For $\xi\in R$ set $\xi' = \xi - \pi(\xi)$.
Specializing $a = X(e_i, \xi' e_j)$, $b = x_{ij}(\pi(\xi))$, $c = X(e_j, \eta'e_k)$, $d = x_{jk}(\pi(\eta))$ and using the definition of the conjugation action 
one computes the commutator $[\psi(x_{ij}(\xi), \psi(x_{jk}(\eta)]$ as follows:
$$\left(X(e_i, \xi' e_j) \cdot X(e_j + \pi(\xi)e_i, \eta'e_k) \cdot X(e_i, \pi(\eta)\xi' e_k - \xi' e_j) \cdot X(e_j, -\eta'e_k),\ x_{ik}(\pi(\xi\eta))\right).$$
After applying additivity relations \eqref{add2}, \eqref{add3} and moving the factor $X(e_i, -\xi' e_j)$ to the left hand side of the formula by means of relation \eqref{conj2}
one gets the following expression
$$(X(e_j, \eta'e_k) \cdot X(e_i, (\xi' \eta' + \pi(\xi)\eta' + \pi(\eta)\xi')e_k \cdot X(e_j, -\eta'e_k),\ x_{ik}(\pi(\xi\eta)).$$
Since $\xi'\eta' + \pi(\xi)\eta' + \pi(\eta)\xi' = \xi\eta - \pi(\xi\eta)$ the above expression clearly simplifies to $\psi(x_{ik}(\xi\eta))$, as claimed.

It remains to check that the map $\varphi$ given by $\varphi\left((X(u, v), 1)\right) = X(u, v),$ $\varphi\left((1, x_{ij}(\xi))\right) = X(e_i, \sigma(\xi)e_j)$
is well-defined and inverse to $\psi$.
\end{proof}

@@

In order to construct the arrow $\kappa$ we use the presentation from \cref{prop:TulPres}.
We set $$\kappa(X(u,v)) = F(u,v),\ u\in \E(n, R)e_1,\ v\in I^n.$$ 
In view of \cref{rk:T123'} it suffices to check that $F(u,v)$ satisfy the relations \eqref{add2}, \eqref{conj2}, \eqref{add3'}.
The validity of relations \eqref{add2}, \eqref{conj2} is obvious while \eqref{add3'} follows from \eqref{add5} and \eqref{coinc}, indeed:
$$ F(e_i, rae_k) F(e_j, ae_k) = S(rae_i, e_k) S(ae_j, e_k) = S(a(re_i+e_j), e_k) = F(e_ir+e_j, ae_k). \qedhere $$

@@

Since $\pi^*(\iota(F(u, v))) = X(\pi(u),\,0)=1$ and $\pi^*(\iota(S(u,v)) = Y(0,\,\pi(v)) = 1$ we get that $\mathrm{Im}(\iota)\subseteq\Ker(\pi^*)=\St(n,\,R,\,I)$. 
Thus, we have demonstrated the following result.

\begin{prop} \label{lm:map-iota}
 For a splitting ideal $I \trianglelefteq R$ and $n\geq 4$ the groups $\St^*(n,\,R,\,I)$ and $\St(n,\,R, I)$ are isomorphic.
% there exists a well-defined homomorphism $$\iota\colon\St^*(n,\,R,\,I)\rightarrow\St(n,\,R, I) \subseteq \St(n, R)$$ 
% satisfying the equalities $\iota (F(u,\,v)) = X(u,\,v)$ and $\iota (S(u,\,v)) = Y(u,\,v)$. 
\end{prop}
%\begin{proof}
%We already know that $\iota$ is well-defined. It suffices to check that the image of $\iota$ is contained in $\St(n,\,R,I)$.
%Denote the canonical projection $R\epi R/I$ by $\rho$.
%Clearly, the induced group morphism $\rho^*\colon\St(n,\,R)\rightarrow\St(n,\,R/I)$ sends $X(u,\,v)$ to $X(\rho(u),\,\rho(v))$ and $Y(u,\,v)$ to $Y(\rho(u),\,\rho(v))$. 
%Observe that $\E(n,\,R)\epi\E(n,\,R/I)$ is surjective, therefore for $x\in\E(n,\,R)e_1$ one has $\rho(x)\in\E(n,\,R/I)e_1$. 
%, as required. \end{proof}

@@

\section{The local-global principle for $\Kt$.} \label{sec:lgp}
%At this point we have demonstrated for $n\geq4$ and a splitting ideal $I$ that $\St^*(n,\,R,\,I)$ and $\St(n,\,R,\,I)$ are isomorphic.
Now we turn to the main result of this paper, namely, we construct the map $$T\colon\St(n,\,R_a[X],\,XR_a[X])\rightarrow\St(n,\,R\ltimes XR_a[X]),\ \text{ for} n\geq4.$$ 
The local-global principle for the Steinberg group and centrality of $\mathrm K_2$ will formally follow from the existence of this map, see~\cite[Lemmas~15--16, and the proof of Theorem~2]{SCh}.

We will prove a somewhat more general result. Let $B$ be a commutative unital ring.
A $B$-module $M$ is called \emph{$r$-divisible} if for every $m\in M$ there exists $m'\in M$ such that $rm' = m$ (we denote such $m'$ by $\frac{m}{a}$).
Clearly, $M$ is $a$-divisible if and only if the morphism of principal localization $\lambda_a\colon M \to M_a$ is an isomorphism.

\begin{tm} \label{a3map} Let $B$ be a ring, $a\in B$ and let $I$ be an ideal $B$ that is $a$-divisible.
Then there exists a map $T\colon\St(n,\,B_a,\,I)\rightarrow\St(n,\,B)$ which makes the following diagram commute.
$$\xymatrix{ \St^*(2n,\,B,\,I)\ar@<-0.0ex>[rr]^{\iota}\ar@<-0.0ex>[d]_{\lambda_{a}^*}&&\St(2n,\,B)\ar@<-0.0ex>[d]^{\lambda_{a}^*}\\
             \St^*(2n,\,B_a,\,I)\ar@<-0.0ex>[rr]_{\iota}\ar@{-->}[rru]_{T}&&\St(2n,\,B_a)}$$ \end{tm}
